### 2015-06-01 inst/doc/summary.rmd
a few summaries & graphs

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Timings}
-->

```{r opts, echo = FALSE, message = FALSE}
library("knitr")
```

```{r loadpkg,message=FALSE}
library("jsonlite")
library("Timings")
library("plyr") ## handy for ddply, rename
library("ggplot2"); theme_set(theme_bw())
```
```{r}
lf <- list.files(system.file("JSON",package="Timings"),full.names=TRUE)
L1 <- lapply(lf,fromJSON)
sfun <- function(x) {
    ff <- x$models$fits[[1]]
    dd <- data.frame(ds=x$dsname,
               ff[c("deviance","optimizer","function","feval")])
    dd$time <- sapply(ff$time,tail,1)
    ## 'function' and 'deviance' both make trouble for NSE
    dd <- rename(dd,replace=c(function.="fun",deviance="dev"))
    dd <- ddply(dd,"ds",transform,
                ddev=dev-min(dev,na.rm=TRUE),
                reltime=time/min(time,na.rm=TRUE))
    dd
}
## structure of data objects differs ...
has.dsnames <- sapply(lapply(L1,names),function(x) "dsname" %in% x)
sdf <- do.call(rbind,lapply(L1[has.dsnames],sfun))
sdf <- transform(sdf,method=paste(fun,optimizer,sep="."))
sdf <- transform(sdf,method=reorder(method,time),
                     ds=reorder(ds,time))
sdf <- sdf[with(sdf,order(method,ds)),]

## timings
## (we've already ordered methods by average time)
ggplot(sdf,aes(time,method,colour=ds))+
    geom_point()+
    geom_path(aes(group=ds))+
    scale_x_log10()

ggplot(sdf,aes(reltime,method,colour=ds))+
    geom_point()+
    geom_path(aes(group=ds))+
    scale_x_log10()

## taking
ggplot(sdf,aes(ddev,method,colour=ds))+
    geom_point()+
        geom_path(aes(group=ds))+
            geom_vline(xintercept=0.01,linetype=2)+
                scale_x_log10()
ggplot(sdf,aes(ddev,ds,colour=method))+
    geom_point()+
        geom_path(aes(group=method))+
            geom_vline(xintercept=0.01,linetype=2)+
                scale_x_log10()

```